Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3\baselineskip}

### 2017-03-21 paper/ms.tex
Added transit refs; described R&P; updated acks

In this example, we inject the signal of a simulated exoplanet transit into a
real \kepler\ light curve and then demonstrate that we can recover the true
physical parameters of the exoplanet while modeling the stellar variability
using \celerite.
This example is different from all the previous examples because in this case,
we are uninterested in the inferred parameters of the covariance model.
Instead, we're interested in inferring constraints on the parameters of the
mean model.
In \eq{gp-likelihood} these parameters are called $\bvec{\theta}$ and in this
example, the mean function $\mu_\bvec{\theta}(t)$ is a limb-darkened transit
light curve \dfmtodo{CITE MA} parameterized by a period $P$, a transit
duration $T$, a phase or epoch $t_0$, an impact parameter $b$, the radius
of the planet in units of the stellar radius $R_P/R_\star$, and several
parameters describing the limb-darkening profile of the star \dfmtodo{CITE}.
As in the previous example, we model the stellar variability using a GP model
with a kernel given by \eq{rot-kernel}.

@@

\dfmtodo{Add discussion of original R\&P method here.}

@@

Further progress was made by parameterizing the covariance matrix with simple,
analytic functions that describe the autocorrelation function.  The parameters
of these functions can then be optimized to best match the observed variability
pattern of the residuals (after subtracting a transit model).  The disadvantage of
this approach is that the functions used are chosen somewhat arbitrarily (e.g.\
the exponential-squared function), and the Cholesky decomposition of the covariance
matrix for computing the likelihood takes ${\cal O}(N^3)$ operations which prohibits
application of this technique to large datasets.

@@

EA acknowledges support from NASA grants NNX13AF20G, NNX13AF62G, and
NASA Astrobiology Institute's Virtual Planetary Laboratory, supported
by NASA under cooperative agreement NNH05ZDA001C.

>>>>

In this example, we inject the signal of a simulated exoplanet transit into a
real \kepler\ light curve and then demonstrate that we can recover the true
physical parameters of the exoplanet while modeling the stellar variability
using \celerite.
This example is different from all the previous examples because in this case,
we are uninterested in the inferred parameters of the covariance model.
Instead, we're interested in inferring constraints on the parameters of the
mean model.
In \eq{gp-likelihood} these parameters are called $\bvec{\theta}$ and in this
example, the mean function $\mu_\bvec{\theta}(t)$ is a limb-darkened transit
light curve \citep{Mandel:2002} parameterized by a period $P$, a transit
duration $T$, a phase or epoch $t_0$, an impact parameter $b$, the radius
of the planet in units of the stellar radius $R_P/R_\star$, and several
parameters describing the limb-darkening profile of the star \citep{Claret:2011}.
As in the previous example, we model the stellar variability using a GP model
with a kernel given by \eq{rot-kernel}.

@@

The \celerite\ model is based upon a method developed by Press \& Rybicki twenty 
years ago \citep{Rybicki:1995}.  They observed that a correlation matrix consisting 
of a single exponential kernel has an inverse that is tri-diagonal;
the coefficients and solution of a tri-diagonal matrix scale in proportion to $O(N)$.
This method does not require an extended matrix, and thus speeds up the solution 
by a factor of $\approx 2$ for large $N$, but 
the requirement of an exponential kernel is extremely restrictive,
equivalent to assuming an Ornstein--Uhlenbeck process, which tends to show sharp 
variations in the light curve, and cannot describe quasi-periodicity.

@@

Further progress was made by parameterizing the covariance matrix with simple,
analytic functions that describe the autocorrelation function \citep{Gibson:2012}.  
The parameters of these functions can then be optimized to best match the observed 
variability pattern of the residuals (after subtracting a transit model).  The 
disadvantage of this approach is that the functions used are chosen somewhat 
arbitrarily (e.g.\ the exponential-squared function), and the Cholesky decomposition 
of the covariance matrix for computing the likelihood takes ${\cal O}(N^3)$ operations 
which prohibits application of this technique to large datasets.

@@

This work was performed in part under contract with the Jet Propulsion Laboratory (JPL) 
funded by NASA through the Sagan Fellowship Program executed by the NASA Exoplanet Science Institute.

EA acknowledges support from NASA grants NNX13AF20G, NNX13A124G, NNX13AF62G, 
from National Science Foundation (NSF) grant AST-1615315, and from
NASA Astrobiology Institute's Virtual Planetary Laboratory, supported
by NASA under cooperative agreement NNH05ZDA001C.
